---
title: "Assignment 2"
author: "Shuo Mao 437681258"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r echo=FALSE, include=FALSE}
library(tidyverse)
```


```{r}
lotto <- read.csv('/Users/vicmon/STATS707_UOA/Assignment_2/luckylotto.csv')
tree <- read.csv('/Users/vicmon/STATS707_UOA/Assignment_2/Treesize.csv')
THC <- read.csv('/Users/vicmon/STATS707_UOA/Assignment_2/THCdata_strain.csv')
```


# It is well known that many people try to improve their chances of winning a lottery by using lucky numbers or buying their tickets from certain lucky stores. To estimate what proportion of population would buy from a lucky store, a question was asked in a class survey. Participants were asked to respond Yes/No to the following question: “If you learned that a certain store has sold the winning lotto ticket a few times in the past year, would you be tempted to buy a ticket from this store?” In the survey, 12 out of 43 respondents said ‘Yes’. These responses are stored in an Excel file luckylotto.CSV on Canvas.


## Task 1:
### a. Based on this data, what will be your estimate for the proportion of people in the whole population that would consider buying their tickets this way? [5 Marks] 

```{r}
# loading data

```


### b. Based on the key concepts of Statistical Inference that you have learned so far, Explain why this estimate can be considered to be a good estimate? [5 Marks]

## Task 2:
### a. Find the confidence intervals for your estimate at the 95% and the 99% levels and interpret each. Specify what the ‘confidence’ actually means and what it doesn’t. [6 Marks]


### b. Explain why the 99% confidence interval is wider than the 95% one? [4 Marks]
